Verification of concurrent design patterns with data

### 2019-04-17 2018-Arch-pNets/Coordination Camera ready/arch-pNets.tex
2 proprietes ajoutees en page 15

\paragraph{Practical experiments.}%\noteSB{J'ai repris ce paragraphe}
%\noteLH{j ai nomme ce paragraphe}
The above encoding provides a mechanism for the symbolic verification
of architectures using our existing tool
\cite{HMZ-FORTE2016QBMZ-AVOCS18} to compute the open automaton
semantics of an open pNet. This tool computes open transitions from
pLTS behaviours and synchronisation vectors of the pNet, then uses an
SMT engine to check satisfiability of their predicates, minimising the
size of the resulting automaton.
%% The encoding into the SMT language involves the definition of the
%% action and data domains and operators, and the axiomatisation of
%% domain-dependant properties, to allow the SMT engine to be as precise
%% as possible (even if it cannot be complete).
%% The resulting open automaton will later be subject to (symbolic)
%% model-checking and equivalence checking algorithms, themselves
%% requiring automatic proof assistance. \todoEM{trop flou pour le
%%   moment, mais je n'ose pas dire des trucs trop precis... meme si on a
%%   deja une implem de la FH-bisim, elle n'est pas validee!}

In \fig{ArchFailure:OA} we show the full open automaton
obtained from the pNet in \fig{FailureTimer:pNet}.  Due to space limitations, we do not show the
details of the open transitions, but only the assignments of state
variables, some useful parts of the predicates, and the resulting
action; full details can be found in 
\app{full-results}.
%
%\noteSB{Removed ``As an example...'', since that \ex{TimerOT3} was
%  simplified and does not correspond to the real OT.}
%% As an example, $OT_3$ is the one
%% explained in \ex{TimerOT3} (page~\pageref{ex:TimerOT3}).
This automaton has 12 transitions, including those encoding various
possible firing of some interactions (e.g. OT 2 and 3 for
{\PortFail}). It looks like the rough behaviour is
reasonable, although a detailed study shows that the property stated
in page~\pageref{property:reset} does not hold, because of the
{\PortFail} loops on states $11$ and $12$. This is because we did not
use the maximal progress assumption here. If we do, we get the
corrected behaviour in \fig{ArchFailure:OA-MaxProgress}
(\app{full-results}), in which $OT_7$ has disappeared, and $OT_2$ and
$OT_{10}$ have been restricted to $b0=\false$. This one verifies all
the properties listed in \app{properties}.

>>>>

\paragraph{Practical experiments.}
The above encoding provides a mechanism for the symbolic verification
of architectures using our existing tool
\cite{HMZ-FORTE2016QBMZ-AVOCS18} to compute the open automaton
semantics of an open pNet. This tool computes open transitions from
pLTS behaviours and synchronisation vectors of the pNet, then uses an
SMT engine to check satisfiability of their predicates, minimising the
size of the resulting automaton.
%% The encoding into the SMT language involves the definition of the
%% action and data domains and operators, and the axiomatisation of
%% domain-dependant properties, to allow the SMT engine to be as precise
%% as possible (even if it cannot be complete).
%% The resulting open automaton will later be subject to (symbolic)
%% model-checking and equivalence checking algorithms, themselves
%% requiring automatic proof assistance. \todoEM{trop flou pour le
%%   moment, mais je n'ose pas dire des trucs trop precis... meme si on a
%%   deja une implem de la FH-bisim, elle n'est pas validee!}

In \fig{ArchFailure:OA} we show the full open automaton
obtained from the pNet in \fig{FailureTimer:pNet}.  Due to space limitations, we do not show the
details of the open transitions, but only the assignments of state
variables, some useful parts of the predicates, and the resulting
action; full details can be found in 
\app{full-results}.
%
%\noteSB{Removed ``As an example...'', since that \ex{TimerOT3} was
%  simplified and does not correspond to the real OT.}
%% As an example, $OT_3$ is the one
%% explained in \ex{TimerOT3} (page~\pageref{ex:TimerOT3}).
This automaton has 12 transitions, including those encoding various
possible firing of some interactions (e.g. OT 2 and 3 for
{\PortFail}).

\add{Model-checking of open-automaton is out of the scope of this paper,
though the resulting automaton here is small enough to observe the
kind of properties we can prove. It is clear that our encoding allows
to test specific values of state-variables in formulas, like e.g.:
  \begin{equation}
    \label{eq:prop:z:top}
    \AW[\quad]{(z = \topinterval)}{\PortFail} \land\quad
    \AG \bigl(
    \PortReset \rightarrow \AW[]{(z = \topinterval)}{\PortFail}
    \bigr).
  \end{equation}
that says that \emph{as long as no failure has occured, the $z$ variable of the
$\NameTimer$ component has the value $\topinterval$}.

But we can also (as long as we get a proper axiomatisation of our data
operators in the SMT engine) handle more involved data properties,
like here the fact that \emph{a reset can only be requested within the
specified delay after a failure:}
  \begin{multline}
    \label{eq:prop:to:delay}
    \forall T_0, T_1 \in \sZ,\
    \AG \Big((\PortFail \land \NameTimer.t = T_0) \rightarrow
    \\
    \AW[]{\bigl((\PortAsk \land \NameTimer.t = T_1) \rightarrow
      (T_1 - T_0 \in \NameCtrl.\mathit{zone})\bigr)
    }{
      (\PortReset \lor \PortResume)
    }\Big),
  \end{multline}
}

Last, a detailed study shows that the safety property stated
in page~\pageref{property:reset} does not hold, because of the
{\PortFail} loops on states $11$ and $12$. This is because we did not
use the maximal progress assumption here. If we do, we get the
corrected behaviour in \fig{ArchFailure:OA-MaxProgress}
(\app{full-results}), in which $OT_7$ has disappeared, and $OT_2$ and
$OT_{10}$ have been restricted to $b0=\false$. This one verifies all
the properties listed in \app{properties}.

\noteEMin{on peut encore bouger deux lignes de l'annexe D.2, pour
  mentionner ici une propriete de liveness, mais ca va dependre de la
  place, c'est deja un peu long ici...}
%% %****************************************************************
